---
title: "SNAX III uptake"
output: pdf_document
---
# SNAX III nutrient uptake

This code is used to estimate nutrient uptake in the experimental streams. We use bootstrapping to account for uncertainty in the conservative concentration of nutrients in each stream and a mass balance approach to estimate the rate of nutrient uptake in the streams. 

Analysis was with R version 3.6.2 (2019-12-12)


```{r echo=FALSE}
#### load packages
rm(mean)
library(ggplot2)
library(bbmle)
library(Rmisc)
library(reshape2)
library(dplyr)
library(cowplot)


set.seed(207)
### load data
setwd("C:\\Users\\nt78066\\OneDrive - University of Georgia\\Documents\\SNAX_3_nutrient_uptake\\")
###
pump.info<-read.csv("all.noon.pump.information_clean.csv" ) # processed in camble.data.logger.info.noon.R and
barrel.conc<-read.csv("conc_nutrient_in_barrels_every_day_clean.csv") #processed in nutrient_conc_in_barrels.R
all.nuts<-read.csv("bottom_of_reach_nutrients.csv")
drip<-read.csv("drip_rate_measurementes.csv")
zero.m.nuts<-read.csv("nutrients_above_0m_clean.csv")
width<-read.csv("snax3_wetted_widths.csv" )
slug<-read.csv("all.salt.slugs.csv")

######### formatting dates
zero.m.nuts$date<-as.Date(strptime(zero.m.nuts$date,format=("%Y-%m-%d")))
pump.info$date<-as.Date(strptime(pump.info$date.time,format=("%Y-%m-%d")))
barrel.conc$date<-as.Date(strptime(barrel.conc$date,format=("%Y-%m-%d")))
all.nuts$Date<-as.Date(strptime(all.nuts$Date,format=("%Y-%m-%d")))
drip$date<-as.Date(strptime(drip$date,format=("%Y-%m-%d")))
width$date<-as.Date(strptime(width$date,format=("%Y-%m-%d")))
slug$date<-as.Date(strptime(slug$Date,format=("%Y-%m-%d")))

```
Some negative water depths in watersheds 7 and 8. Watershed 7 is dealt with separately,
discarding some bad looking data from the depth logger in WS08
```{r}
pump.info<-pump.info[pump.info$WaterDep>0,]

drip<-drip[drip$flow.rate.ml.stroke>0,]  ## The zeros are from days that the pumps were not moving, so the 
## number of strokes is zero (thus, if the # of strokes is zero the product of # of strokes x stroke volume remains zero regardless of flow rate). Importantly, we don't want to sample these zero values when selecting values for days that the were pumps were pumping. The pumps were programmed to shut off at extreme high flows. The frequency of these events can be seen in appendix figures S1 and S2 as loadings near zero that have no associated error.  
```


Getting background nutrient data sorted correctly.
```{r}
zero.m.nuts2<-data.frame(date=zero.m.nuts$date,stream=zero.m.nuts$stream,background.DIN=zero.m.nuts$DIN..µg.L.,background.P=zero.m.nuts$SRP..µg.L.)

names(all.nuts)[names(all.nuts) == "Date"] <- "date"
names(all.nuts)[names(all.nuts) == "Stream"] <- "stream"


all.nuts<-merge(all.nuts,zero.m.nuts2,by=c("date","stream"))

```

#rating curves for each stream

```{r}

streams_with_good_curves<-c("WS06","WS08","WS09","WS10")
NP.list<-c("32:1","16:1","2:1","8:1")
slug$discharge<-slug$Measured.Q..L.s.1.

for (i in 1:4){
  
stream<-streams_with_good_curves[i]

power_law<-lm(log(discharge)~log(Stage.meters),slug[slug$Stream==stream,])
spl<-summary(power_law)

model.fit<-data.frame(stage=seq(0,max(slug[slug$Stream==stream,"Stage.meters"],na.rm=TRUE),by=0.001))
model.fit$discharge<-exp(coef(power_law)[2]*log(model.fit$stage)+coef(power_law)[1])

rating.curve.plot<-ggplot(slug[slug$Stream==stream,],aes(x=Stage.meters,y=discharge))+theme_classic()+geom_point()+ylab("Discharge L/s")+xlab("Stage height (m)")+annotate("text",0.04,20,label=paste("R2 = ",round(spl$r.squared,2)),size=6)+annotate("text",0.04,17,label=paste("Target N:P=",NP.list[i]),size=6) +geom_line(data=model.fit,aes(x=stage,y=discharge)) +theme(text =element_text(size=15))

tiff(filename=paste(stream,"_discharge",".tiff",sep=""),units="in",res=800,width=6,height=6,compression="lzw")
print(rating.curve.plot)
dev.off()


}
  
  

```



# Uncertaintiy

If there is no uptake then the nutrient concentration at the bottom of the reach is a combination of the background nutrients and the nutrients which were added to the stream from the pump system.There are some sources of uncertainty in these nutrient concentrations.

1.) The volume of nutrient solution moved per turn of the pump are somewhat variable. We have measurements of these volumes, but not on every date. 
2.) There is some uncertainty in the flow-discharge relationship (stage discharge relationship)
3.) There is some uncertainty in the relationship between discharge and stream width



## Functions

Functions to estimate stage-discharge relationships, discharge-width relationships, and conservative concentrations of DIN and SRP

```{r}
# The quality of the rating curves varies across the stream, by sampling from the curves (with replacement) and recalculating the relationships our analysis propagates this uncertainty into our estimates of nutrient uptake from the streams where we are less certain about the flow stage relationships.

setwd("C:\\Users\\nt78066\\OneDrive - University of Georgia\\Documents\\SNAX_3_nutrient_uptake\\")
pump.info$corrected_discharge<-NA


colnames(slug)[colnames(slug)=="Measured.Q..L.s.1."] <- "discharge"


# this function takes the stream identify, subsets the data for that stream, fits a power law model to a subset of the data then predicts the data on each day where stage measurements exist

recalculate_discharge<-function(stream_id){
  c2<-slug[slug$Stream==paste(stream_id),]
  flow.mod2<-lm(log(discharge)~log(Stage.meters),data=c2[sample(nrow(c2), nrow(c2),replace=TRUE), ])
  
  
  corrected.flow<-exp(coef(flow.mod2)[2]*log(pump.info[pump.info$stream==paste(stream_id),"WaterDep"])+coef(flow.mod2)[1])
  return(data.frame(corrected.flow))
  
  
}

######### WS 7 stage data has some issues, using correlations between slugs to convert

slug2<-melt(slug,measure.vars = "discharge",id.vars=c("Date","Stream"))

slug3<-dcast(Date~variable+Stream,data=slug2,mean)

## stream 9 has a record that looks good, going to estiamte stream 7 flow based ona regression

stream_7_model<-lm(discharge_WS07~discharge_WS09+discharge_WS06,data=slug3)


discharge1<-ggplot(slug3,aes(y=discharge_WS07,x=discharge_WS09))+theme_classic()+geom_smooth(method="lm",se=FALSE,size=3)+xlab("Discharge (L/s, stream N:P=2)")+ylab("Discharge (L/s, stream N:P=128)")+geom_point(size=3)+theme(text = element_text(size=20))

discharge2<-ggplot(slug3,aes(y=discharge_WS07,x=discharge_WS06))+theme_classic()+geom_smooth(method="lm",se=FALSE,size=3)+xlab("Discharge (L/s, stream N:P=32)")+ylab("Discharge (L/s, stream N:P=128)")+geom_point(size=3)+theme(text = element_text(size=20))

tiff(filename="discharge_correction.tiff",units="in",res=800,width=6,height=10,compression="lzw")
multiplot(discharge1,discharge2)
dev.off()


recalculate_discharge_stream7<-function(){
  
  
  WS6<-data.frame(date=e$date,discharge_WS06=e$corrected.flow)
  WS09<-data.frame(date=b$date,discharge_WS09=b$corrected.flow)
  
  temp<-inner_join(WS6,WS09,by="date")
  temp<-temp[complete.cases(temp),]
  
  corrected.flow<-coef(stream_7_model)[1]+temp$discharge_WS09*coef(stream_7_model)[2]+temp$discharge_WS06*coef(stream_7_model)[3]
  
  min.flow<-NA
  max.flow<-NA
  
  
  output<-data.frame(date=temp$date,corrected.flow,min.flow,max.flow)
  
  return(output)
  
  
}


# Similarly need to incorperate variation in the relationship between flow and width. 

width.model<-function(stream_id){
  sub.w<-width2[width2$stream==stream_id,]
  width.mod2<-lm(log(wetted.width.m)~log(corrected.flow),data=sub.w[sample(nrow(sub.w), nrow(sub.w),replace=TRUE), ])
  
  exp(coef(width.mod2)[1])*pump.info2[pump.info2$stream==paste(stream_id),"corrected.flow"]**coef(width.mod2)[2]
  
  }

conservative_concentration_DIN<-function(stream_id){
  # this calculate the theoretical concentraiton of DIN as the background concentration plus the input rate times the portion of inputs that happened upstream of the sample site
  all.data[all.data$stream==paste(stream_id),"background.DIN"]+all.data[all.data$stream==paste(stream_id),"input.rate.din"]
}

conservative_concentration_SRP<-function(stream_id){
  # this calculate the theoretical concentraiton of SRP as the background concentration plus the input rate times the portion of inputs that happened upstream of the sample site
  all.data[all.data$stream==paste(stream_id),"background.P"]+all.data[all.data$stream==paste(stream_id),"input.rate.srp"]
  }



```

# Calculating theoretical - no uptake - concentrations of nutrients

```{r,echo=FALSE}
inputs1<-merge(barrel.conc,drip,by=c("stream","date"),all.x=TRUE)

compiled_data<-data.frame()
pump.info$stream.width<-NA

daily.loading<-data.frame()

#### Looping through 1000 iterations in each stream
for (i in 1:1000){
stroke_vector<-inputs1$flow.rate.ml.stroke ## this gets reinitialized every rep

  #select random value of stroke rate when there isn't one on the day of measurement 
for (m in 1:nrow(inputs1)){if(is.na(stroke_vector[m])==TRUE){
  stream.meas<-drip[drip$stream==inputs1$stream[m],]
  stroke_vector[m]<-base::sample(stream.meas$flow.rate.ml.stroke,1)
  
  
  
}}

inputs1$stroke.rate.ml.stroke.complete<-stroke_vector

a<-cbind(pump.info[pump.info$stream=="WS10",],recalculate_discharge("WS10"))
b<-cbind(pump.info[pump.info$stream=="WS09",],recalculate_discharge("WS09"))
c<-cbind(pump.info[pump.info$stream=="WS08",],recalculate_discharge("WS08"))
e<-cbind(pump.info[pump.info$stream=="WS06",],recalculate_discharge("WS06"))
discharge_object<-recalculate_discharge_stream7()
d<-left_join(pump.info[pump.info$stream=="WS07",],discharge_object,by="date")


pump.info2<-bind_rows(a,b,c,d,e)
pump.info2$stream<-as.factor(pump.info2$stream)

width2<-merge(width,pump.info2,by=c("stream","date"))

pump.info2<-pump.info2[is.na(pump.info2$stream)==FALSE,]

pump.info2[pump.info2$stream=="WS10","stream.width"]<-width.model("WS10")
pump.info2[pump.info2$stream=="WS09","stream.width"]<-width.model("WS09")
pump.info2[pump.info2$stream=="WS08","stream.width"]<-width.model("WS08")
pump.info2[pump.info2$stream=="WS07","stream.width"]<-width.model("WS07")
pump.info2[pump.info2$stream=="WS06","stream.width"]<-width.model("WS06")

#check
#ggplot(pump.info,aes(x=corrected.flow,y=stream.width,color=stream))+geom_point()

inputs2<-merge(inputs1,pump.info2,by=c("stream","date"))



inputs2$input.rate.srp<-(((inputs2$number.of.strokes*inputs2$stroke.rate.ml.stroke.complete)/(15*60))*
                           (inputs2$conc.srp.in.barrel*1000))/inputs2$corrected.flow
# 1000 is a unit conversion for the concentrations
# dividing by 15*60 converts the number of strokes from the number per 15 mins to the number per S

inputs2$input.rate.din<-((inputs2$number.of.strokes*inputs2$stroke.rate.ml.stroke.complete)/(15*60)*
                           (inputs2$conc.DIN.in.barrel*1000))/inputs2$corrected.flow

## input rate of SRP and DIN are in the units of ug N or P/ L water
inputs2$iteration<-i

daily.loading<-rbind(daily.loading,inputs2)

all.data<-merge(all.nuts,inputs2,by=c("stream","date"))



all.data$conservative_DIN<-NA
all.data$conservative_SRP<-NA

# WS 06
all.data[all.data$stream=="WS06","conservative_DIN"]<-conservative_concentration_DIN("WS06")
all.data[all.data$stream=="WS06","conservative_SRP"]<-conservative_concentration_SRP("WS06")

# WS 07
all.data[all.data$stream=="WS07","conservative_DIN"]<-conservative_concentration_DIN("WS07")
all.data[all.data$stream=="WS07","conservative_SRP"]<-conservative_concentration_SRP("WS07")

# WS 08
all.data[all.data$stream=="WS08","conservative_DIN"]<-conservative_concentration_DIN("WS08")
all.data[all.data$stream=="WS08","conservative_SRP"]<-conservative_concentration_SRP("WS08")

# WS 09
all.data[all.data$stream=="WS09","conservative_DIN"]<-conservative_concentration_DIN("WS09")
all.data[all.data$stream=="WS09","conservative_SRP"]<-conservative_concentration_SRP("WS09")

# WS 10
all.data[all.data$stream=="WS10","conservative_DIN"]<-conservative_concentration_DIN("WS10")
all.data[all.data$stream=="WS10","conservative_SRP"]<-conservative_concentration_SRP("WS10")


compiled_data<-rbind(compiled_data,all.data)
}
```
NaNs warning occurs when there is no measured background concentration of nutrients - these should not be removed because there are cases where there is a measurement for N and not P or the opposite


```{r}
### graphing loading during the experiment for appendix S1
p.input.sum<-aggregate(input.rate.srp~date+stream,daily.loading,quantile,probs=c(0.025,0.5,0.975))

mean.background.conc.p<-aggregate(background.P~stream,data=all.nuts,mean)

p.input.sum2<-merge(p.input.sum,mean.background.conc.p,by="stream")
p.input.sum2$median<-p.input.sum2$input.rate.srp[,2]+p.input.sum2$background.P
p.input.sum2$lower<-p.input.sum2$input.rate.srp[,1]+p.input.sum2$background.P
p.input.sum2$upper<-p.input.sum2$input.rate.srp[,3]+p.input.sum2$background.P


st<-"WS06"

stream_6_loading<-ggplot(p.input.sum2[p.input.sum2$stream==st,],aes(x=date,y=median))+geom_point(color="red")+geom_errorbar(aes(ymin=lower,ymax=upper,color="red"),width=0.000001,alpha=0.5)+theme_classic()+geom_point(data=all.nuts[all.nuts$stream==st,],aes(x=date,y=PO4.P),color="black")+theme(text =element_text(size=15))+theme(legend.position = "none")+ylab("")#+geom_line(data=daily.loading[daily.loading$iteration==1 & daily.loading$stream==st,],aes(x=date,y=corrected.flow*2),alpha=0.5)

st<-"WS07"

stream_7_loading<-ggplot(p.input.sum2[p.input.sum2$stream==st,],aes(x=date,y=median))+geom_point(color="red")+geom_errorbar(aes(ymin=lower,ymax=upper,color="red"),width=0.000001,alpha=0.5)+theme_classic()+geom_point(data=all.nuts[all.nuts$stream==st,],aes(x=date,y=PO4.P),color="black")+theme(legend.position = "none")+theme(text =element_text(size=15))+ylab("")#+geom_line(data=daily.loading[daily.loading$iteration==1 & daily.loading$stream==st,],aes(x=date,y=corrected.flow*2),alpha=0.5)

st<-"WS08"

stream_8_loading<-ggplot(p.input.sum2[p.input.sum2$stream==st,],aes(x=date,y=median))+geom_point(color="red")+geom_errorbar(aes(ymin=lower,ymax=upper,color="red"),width=0.000001,alpha=0.5)+theme_classic()+geom_point(data=all.nuts[all.nuts$stream==st,],aes(x=date,y=PO4.P),color="black")+theme(legend.position = "none")+theme(text =element_text(size=15))+ylab("Phosphorus Concentration (ug/l)")#+geom_line(data=daily.loading[daily.loading$iteration==1 & daily.loading$stream==st,],aes(x=date,y=corrected.flow*2),alpha=0.5)

st<-"WS09"

stream_9_loading<-ggplot(p.input.sum2[p.input.sum2$stream==st,],aes(x=date,y=median))+geom_point(color="red")+geom_errorbar(aes(ymin=lower,ymax=upper,color="red"),width=0.000001,alpha=0.5)+theme_classic()+geom_point(data=all.nuts[all.nuts$stream==st,],aes(x=date,y=PO4.P),color="black")+theme(legend.position = "none")+theme(text =element_text(size=15))+ylab("")#+geom_line(data=daily.loading[daily.loading$iteration==1 & daily.loading$stream==st,],aes(x=date,y=corrected.flow*2),alpha=0.5)

st<-"WS10"

stream_10_loading<-ggplot(p.input.sum2[p.input.sum2$stream==st,],aes(x=date,y=median))+geom_point(color="red")+geom_errorbar(aes(ymin=lower,ymax=upper,color="red"),width=0.000001,alpha=0.5)+theme_classic()+geom_point(data=all.nuts[all.nuts$stream==st,],aes(x=date,y=PO4.P),color="black")+theme(legend.position = "none")+theme(text =element_text(size=15))+ylab("")#+geom_line(data=daily.loading[daily.loading$iteration==1 & daily.loading$stream==st,],aes(x=date,y=corrected.flow*2),alpha=0.5)

labels<-c("N:P=2","N:P=8","N:P=16","N:P=32","N:P=128")

tiff(filename="./phosphorus.loading.tiff",units="in",res=800,width=10,height=15,compression="lzw")
plot_grid(stream_9_loading,stream_10_loading,stream_8_loading,stream_6_loading,stream_7_loading,labels=labels,ncol=1,label_x=0.75,label_y=0.9,label_size = 20)
dev.off()


### DIN
n.input.sum<-aggregate(input.rate.din~date+stream,daily.loading,quantile,probs=c(0.025,0.5,0.975))

mean.background.conc.n<-aggregate(background.DIN~stream,data=all.nuts,mean)

n.input.sum2<-merge(n.input.sum,mean.background.conc.n,by="stream")
n.input.sum2$median<-n.input.sum2$input.rate.din[,2]+n.input.sum2$background.DIN
n.input.sum2$lower<-n.input.sum2$input.rate.din[,1]+n.input.sum2$background.DIN
n.input.sum2$upper<-n.input.sum2$input.rate.din[,3]+n.input.sum2$background.DIN


st<-"WS06"

stream_6_loading<-ggplot(n.input.sum2[p.input.sum2$stream==st,],aes(x=date,y=median))+geom_point(color="red")+geom_errorbar(aes(ymin=lower,ymax=upper,color="red"),width=0.000001,alpha=0.5)+theme_classic()+geom_point(data=all.nuts[all.nuts$stream==st,],aes(x=date,y=DIN),color="black")+theme(text =element_text(size=15))+theme(legend.position = "none")+ylab("")#+geom_line(data=daily.loading[daily.loading$iteration==1 & daily.loading$stream==st,],aes(x=date,y=corrected.flow*2),alpha=0.5)

st<-"WS07"

stream_7_loading<-ggplot(n.input.sum2[p.input.sum2$stream==st,],aes(x=date,y=median))+geom_point(color="red")+geom_errorbar(aes(ymin=lower,ymax=upper,color="red"),width=0.000001,alpha=0.5)+theme_classic()+geom_point(data=all.nuts[all.nuts$stream==st,],aes(x=date,y=DIN),color="black")+theme(legend.position = "none")+theme(text =element_text(size=15))+ylab("")#+geom_line(data=daily.loading[daily.loading$iteration==1 & daily.loading$stream==st,],aes(x=date,y=corrected.flow*2),alpha=0.5)

st<-"WS08"

stream_8_loading<-ggplot(n.input.sum2[p.input.sum2$stream==st,],aes(x=date,y=median))+geom_point(color="red")+geom_errorbar(aes(ymin=lower,ymax=upper,color="red"),width=0.000001,alpha=0.5)+theme_classic()+geom_point(data=all.nuts[all.nuts$stream==st,],aes(x=date,y=DIN),color="black")+theme(legend.position = "none")+theme(text =element_text(size=15))+ylab("DIN Concentration (ug/l)")#+geom_line(data=daily.loading[daily.loading$iteration==1 & daily.loading$stream==st,],aes(x=date,y=corrected.flow*2),alpha=0.5)

st<-"WS09"

stream_9_loading<-ggplot(n.input.sum2[p.input.sum2$stream==st,],aes(x=date,y=median))+geom_point(color="red")+geom_errorbar(aes(ymin=lower,ymax=upper,color="red"),width=0.000001,alpha=0.5)+theme_classic()+geom_point(data=all.nuts[all.nuts$stream==st,],aes(x=date,y=DIN),color="black")+theme(legend.position = "none")+theme(text =element_text(size=15))+ylab("")#+geom_line(data=daily.loading[daily.loading$iteration==1 & daily.loading$stream==st,],aes(x=date,y=corrected.flow*2),alpha=0.5)

st<-"WS10"

stream_10_loading<-ggplot(n.input.sum2[p.input.sum2$stream==st,],aes(x=date,y=median))+geom_point(color="red")+geom_errorbar(aes(ymin=lower,ymax=upper,color="red"),width=0.000001,alpha=0.5)+theme_classic()+geom_point(data=all.nuts[all.nuts$stream==st,],aes(x=date,y=DIN),color="black")+theme(legend.position = "none")+theme(text =element_text(size=15))+ylab("")#+geom_line(data=daily.loading[daily.loading$iteration==1 & daily.loading$stream==st,],aes(x=date,y=corrected.flow*2),alpha=0.5)

labels<-c("N:P=2","N:P=8","N:P=16","N:P=32","N:P=128")

tiff(filename="./nitrogen.loading.tiff",units="in",res=800,width=10,height=15,compression="lzw")
plot_grid(stream_9_loading,stream_10_loading,stream_8_loading,stream_6_loading,stream_7_loading,labels=labels,ncol=1,label_x=0.75,label_y=0.9,label_size = 20)
dev.off()

```





#Estimating uptake rates

```{r}
compiled_data$p.uptake.ug.m2.sec<-(compiled_data$conservative_SRP-compiled_data$PO4.P)*(compiled_data$corrected.flow)/(compiled_data$stream.width*compiled_data$Meter)

compiled_data$n.uptake.ug.m2.sec<-(compiled_data$conservative_DIN-compiled_data$DIN)*(compiled_data$corrected.flow)/(compiled_data$stream.width*compiled_data$Meter)

```

### Figure 1
One example plot from a stream

```{r}
ex<-compiled_data[compiled_data$NP==16 & compiled_data$date=="2012-07-01" & compiled_data$Meter==70,]

low<-quantile(ex$conservative_SRP,probs=0.025,na.rm=TRUE)
high<-quantile(ex$conservative_SRP,probs=0.975,na.rm=TRUE)

cbbPalette <- c( "#E69F00", "#56B4E9", "#009E73")


example_plot<-ggplot(ex,aes(conservative_SRP))+xlab(expression("Nutrient Concentration ("*mu*g~L^-1*")"))+ scale_y_continuous(limits = c(0,0.043), expand = c(0, 0))+theme_classic()+ylab("Likelihood of conservative concentration")+scale_x_continuous(limits = c(20,125), expand = c(0, 1))+annotate("rect", xmin = 20, xmax = low, ymin = 0, ymax = 0.043,alpha = 0.5,fill="#56B4E9")+annotate("rect", xmin = low, xmax = high, ymin = 0, ymax = 0.043,alpha = 0.5,fill="#E69F00")+annotate("rect", xmin = high, xmax = 115, ymin = 0, ymax = 0.043,alpha = 0.5,fill="#009E73")+geom_density(size=2)+theme(text = element_text(size=15))+annotate("text",x=low-8,y=0.012,label="Net Uptake",size=6)+annotate("text",x=57,y=0.012,label="Balanced/ No Uptake",size=6)+annotate("text",x=high+10,y=0.012,label="Net Release",size=6)+ geom_segment(aes(x = high, y = 0.008, xend = 115, yend = 0.008), arrow = arrow(length = unit(0.5, "cm")),size=2)+geom_segment(aes(x = low, y = 0.008, xend = 20, yend = 0.008), arrow = arrow(length = unit(0.5, "cm")),size=2)+geom_segment(aes(x = low, y = 0.008, xend = high, yend = 0.008),size=2,linetype="longdash")+geom_segment(aes(x = low, y = 0.006, xend = low, yend = 0.01), size=2)+geom_segment(aes(x = high, y = 0.006, xend = high, yend = 0.01), size=2)
example_plot

tiff(filename="example_of_conservative_conc.tiff",units="in",res=800,width=7,height=7,compression="lzw")
example_plot
dev.off()
```



#Is there net uptake on average? (data in table A2)

```{r}
### appendix table values come from here

compiled_data$n.uptake.uM.m2.sec<-compiled_data$n.uptake.ug.m2.sec/14 #convert to molar

N.first<-aggregate( n.uptake.uM.m2.sec~ NP+date,data=compiled_data[compiled_data$Meter==70,], 
          FUN = median )



compiled_data$p.uptake.uM.m2.sec<-compiled_data$p.uptake.ug.m2.sec/31

P.first<-aggregate( p.uptake.uM.m2.sec~ NP+date,data=compiled_data[compiled_data$Meter==70,], 
          FUN = median )
summarySE(P.first,measurevar="p.uptake.uM.m2.sec",groupvars="NP")

aggregate( p.uptake.uM.m2.sec~ NP,data=P.first, 
          FUN = 'quantile', probs=c(2.5, 50, 97.5)/100 )

aggregate( n.uptake.uM.m2.sec~ NP,data=N.first, 
          FUN = 'quantile', probs=c(2.5, 50, 97.5)/100 )

```

#Figure 2
distribution of uptake rates
```{r}
N.quants<-aggregate( n.uptake.ug.m2.sec~ NP+date,data=compiled_data, 
          FUN = 'quantile', probs=c(2.5, 50, 97.5)/100 )

N.quants$n.status<-"balanced"
N.quants[N.quants$n.uptake.ug.m2.sec[,1]>0,"n.status"]<-"uptake"
N.quants[N.quants$n.uptake.ug.m2.sec[,3]<0,"n.status"]<-"release"

N.quants$binary<-0
N.quants[N.quants$n.uptake.ug.m2.sec[,1]>0,"binary"]<-1
N.quants[N.quants$n.uptake.ug.m2.sec[,3]<0,"binary"]<-(-1)

N.quants$nutrient<-"DIN"


####
cbbPalette <- c( "#009E73","#E69F00","#56B4E9")

N.quants$n.status<-factor(N.quants$n.status,levels=c("release","balanced","uptake"))
  

dist.plot.N<-ggplot(N.quants,aes(x=as.factor(NP),fill=n.status))+geom_bar(stat="count",position="fill")+
  theme_classic()+scale_fill_manual(values=alpha(cbbPalette,0.7))+
  theme(text = element_text(size=20))+xlab("")+ylab("Proportion of Observations")+
  guides(fill=guide_legend(title=""))+annotate("text",x=6,y=0.92,label="A  ",size=10)


P.quants<-aggregate( p.uptake.ug.m2.sec~ NP+date,data=compiled_data, 
          FUN = 'quantile', probs=c(2.5, 50, 97.5)/100 )

P.quants$p.status<-"balanced"
P.quants[P.quants$p.uptake.ug.m2.sec[,1]>0,"p.status"]<-"uptake"
P.quants[P.quants$p.uptake.ug.m2.sec[,3]<0,"p.status"]<-"release"


P.quants$p.status<-factor(P.quants$p.status,levels=c("release","balanced","uptake"))

dist.plot.P<-ggplot(P.quants,aes(x=as.factor(NP),fill=p.status))+geom_bar(stat="count",position = "fill")+
  theme_classic()+scale_fill_manual(values=alpha(cbbPalette,0.7))+
  theme(text = element_text(size=20))+xlab("Target N:P ratio")+ylab("Proportion of Observations")+annotate("text",x=6,y=0.92,label="B  ",size=10)
  
multiplot(dist.plot.N,dist.plot.P,cols=1)

tiff(filename="distribution.jan.2022.tiff",units="in",res=600,width=6,height=10,compression="lzw")
multiplot(dist.plot.N,dist.plot.P,cols=1)
dev.off()

#frequencies of different states (reported in results section)
summary(as.factor(P.quants$p.status))

#uptake
78/175 ### 45% uptake
#release
41/175 ### 23% release

#DIN
summary(as.factor(N.quants$n.status))
 
78/167 ### 47% uptake

50/167 ### 30 % release
```

#Figure 3
time series of uptake and release
```{r}
cbbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73","#CC79A7")

P.quants$binary<-0
P.quants[P.quants$p.uptake.ug.m2.sec[,1]>0,"binary"]<-1
P.quants[P.quants$p.uptake.ug.m2.sec[,3]<0,"binary"]<-(-1)
P.quants$nutrient<-"SRP"

P.quants$binary<-P.quants$binary+0.1
#quick output
output<-data.frame(stream=P.quants$NP)


P.quants$NP<-as.factor(P.quants$NP)



plot.data<-rbind(P.quants,setNames(N.quants,names(P.quants)))

patterns_2<-ggplot(plot.data[plot.data$NP==2,],aes(x=date,y=binary,color=nutrient,shape=nutrient))+
  #geom_point(position=position_jitter(width=0, height=0.05))+
  geom_point(size=3)+theme_classic()+scale_color_manual(values=c("#56B4E9",  "#009E73"))+
  scale_y_continuous(breaks = -1:1, labels=c("Net Release","Balanced","Net Uptake"))+ylab(NULL)+theme(axis.text=element_text(size=12))+xlab("")+annotate("text",x=as.Date("2011-07-26"),y=-0.5,label= "N:P = 2",size=6,hjust=0)+xlim(as.Date("2011-07-25"),as.Date("2013-08-22"))+ theme(legend.position = "none")+scale_shape_manual(values=c(2,19))



patterns_8<-ggplot(plot.data[plot.data$NP==8,],aes(x=date,y=binary,color=nutrient,shape=nutrient))+
  #geom_point(position=position_jitter(width=0, height=0.05))+
  geom_point(size=3)+theme_classic()+scale_color_manual(values=c("#56B4E9",  "#009E73"))+
  scale_y_continuous(breaks = -1:1, labels=c("Net Release","Balanced","Net Uptake"))+ylab(NULL)+theme(axis.text=element_text(size=12))+xlab("")+annotate("text",x=as.Date("2011-07-26"),y=-0.5,label= "N:P = 8",size=6,hjust=0)+xlim(as.Date("2011-07-25"),as.Date("2013-08-22"))+ theme(legend.position = "none")+scale_shape_manual(values=c(2,19))
  


patterns_16<-ggplot(plot.data[plot.data$NP==16,],aes(x=date,y=binary,color=nutrient,shape=nutrient))+
  #geom_point(position=position_jitter(width=0, height=0.05))+
  geom_point(size=3)+theme_classic()+scale_color_manual(values=c("#56B4E9",  "#009E73"))+
  scale_y_continuous(breaks = -1:1, labels=c("Net Release","Balanced","Net Uptake"))+ylab(NULL)+theme(axis.text=element_text(size=12))+xlab("")+annotate("text",x=as.Date("2011-07-26"),y=-0.5,label= "N:P = 16",size=6,hjust=0)+xlim(as.Date("2011-07-25"),as.Date("2013-08-22"))+ theme(legend.position = c(0.06,0.86))+scale_shape_manual(values=c(2,19))
  


patterns_32<-ggplot(plot.data[plot.data$NP==32,],aes(x=date,y=binary,color=nutrient,shape=nutrient))+
  #geom_point(position=position_jitter(width=0, height=0.05))+
  geom_point(size=3)+theme_classic()+scale_color_manual(values=c("#56B4E9",  "#009E73"))+
  scale_y_continuous(breaks = -1:1, labels=c("Net Release","Balanced","Net Uptake"))+ylab(NULL)+theme(axis.text=element_text(size=12))+xlab("")+annotate("text",x=as.Date("2011-07-26"),y=-0.5,label= "N:P = 32",size=6,hjust=0)+xlim(as.Date("2011-07-25"),as.Date("2013-08-22"))+ theme(legend.position = "none")+scale_shape_manual(values=c(2,19))
  


patterns_128<-ggplot(plot.data[plot.data$NP==128,],aes(x=date,y=binary,color=nutrient,shape=nutrient))+
  #geom_point(position=position_jitter(width=0, height=0.05))+
  geom_point(size=3)+theme_classic()+scale_color_manual(values=c("#56B4E9",  "#009E73"))+
  scale_y_continuous(breaks = -1:1, labels=c("Net Release","Balanced","Net Uptake"))+ylab(NULL)+theme(axis.text=element_text(size=12))+xlab("")+annotate("text",x=as.Date("2011-07-26"),y=-0.5,label= "N:P = 128",size=6,hjust=0)+xlab("Date")+xlim(as.Date("2011-07-25"),as.Date("2013-08-22"))+ theme(legend.position = "none")+scale_shape_manual(values=c(2,19))


tiff(filename="uptake_release_new_jan2022.tiff",units="in",res=800,width=8,height=10,compression="lzw")
plot_grid(patterns_2,patterns_8,patterns_16,patterns_32,patterns_128,ncol=1,labels="AUTO",label_x=0.95,label_y=0.9)
dev.off()



```


#Figure 4
stoich plots and the associated statistics



```{r}

library(lme4)
library(lmerTest)

stoich<-merge(N.quants,P.quants,by=c("date","NP"),)

bottom<-compiled_data[compiled_data$input.rate.din!=0,]
bottom$conservative.NP<-(bottom$conservative_DIN/14)/(bottom$conservative_SRP/31)

bottom$uptake.NP<-(bottom$n.uptake.ug.m2.sec/14)/(bottom$p.uptake.ug.m2.sec/31)

inputs.stoich<-aggregate(conservative.NP~ NP+date,data=bottom, 
           FUN = 'quantile', probs=c(2.5, 50, 97.5)/100 )

bottom.stoich<-aggregate(uptake.NP~ NP+date,data=bottom, 
           FUN = 'quantile', probs=c(2.5, 50, 97.5)/100 )

stoich<-merge(stoich,inputs.stoich,by=c("date","NP"))
stoich<-merge(stoich,bottom.stoich,by=c("date","NP"))

stoich2<-stoich[stoich$n.status=="uptake" & stoich$p.status=="uptake",]

release<-stoich[stoich$n.status=="release" & stoich$p.status=="release",]


cbbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73","#CC79A7")


stoich2$Target_NP_ratio<-as.factor(stoich2$NP)

se <- function(x) sd(x) / sqrt(length(x)) 

enrichment.conc<-aggregate(conservative.NP[,2]~NP,data=stoich2,mean)
enrichment.conc.se<-aggregate(conservative.NP[,2]~NP,data=stoich2,se)

uptake.ratio<-aggregate(uptake.NP[,2]~NP,data=stoich2,mean)
uptake.ratio.se<-aggregate(uptake.NP[,2]~NP,data=stoich2,se)

stoich3<-data.frame(NP=enrichment.conc[,1],mean.enrichment.ratio=enrichment.conc[,2],enrichment.ratio.se=enrichment.conc.se[,2],uptake.ratio.mean=uptake.ratio[,2],uptake.ratio.se=uptake.ratio.se[,2])

#log.normal.plot<-ggplot(stoich3,aes(x=mean.enrichment.ratio,y=uptake.ratio.mean,color=as.factor(NP),shape=as.factor(NP)))+geom_abline(intercept=0.11158,slope=0.97722,size=1,color="black")+ theme_classic()+ geom_abline(intercept=0,slope=1,size=1,linetype="dashed")+theme(text =element_text(size=20))+ scale_x_log10()+scale_y_log10()+scale_color_manual(values=cbbPalette)+xlab("Added + Background N:P ratio")+ylab("Net uptake N:P ratio")+scale_shape_manual(values=c(0, 1, 2,18,19))+theme(legend.position = c(0.2, 0.81))+xlim(1,166)+ylim(1,166)+geom_point(size=5)+geom_errorbar(aes(ymin=uptake.ratio.mean-uptake.ratio.se*1.96,ymax=uptake.ratio.mean+uptake.ratio.se*1.96),width=0.01)+geom_errorbarh(aes(xmin=mean.enrichment.ratio-enrichment.ratio.se*1.96,xmax=mean.enrichment.ratio+enrichment.ratio.se*1.96),width=0.01)

stoich2$uptake.NP.plot<-stoich2$uptake.NP[,2]
stoich2$conservative.NP.plot<-stoich2$conservative.NP[,2]

log.normal.plot<-ggplot(stoich2,aes(x=conservative.NP.plot,y=uptake.NP.plot,color=as.factor(NP),shape=as.factor(NP)))+geom_abline(intercept=0.11158,slope=0.97722,size=1,color="black")+ theme_classic()+ geom_abline(intercept=0,slope=1,size=1,linetype="dashed")+theme(text =element_text(size=20))+ scale_x_log10()+scale_y_log10()+scale_color_manual(values=cbbPalette)+xlab("Added + Background N:P ratio")+ylab("Net uptake N:P ratio")+scale_shape_manual(values=c(0, 1, 2,18,19))+theme(legend.position = c(0.2, 0.81))+xlim(1,166)+ylim(1,166)+geom_point(size=5)#+geom_errorbar(aes(ymin=uptake.ratio.mean-uptake.ratio.se*1.96,ymax=uptake.ratio.mean+uptake.ratio.se*1.96),width=0.01)+geom_errorbarh(aes(xmin=mean.enrichment.ratio-enrichment.ratio.se*1.96,xmax=mean.enrichment.ratio+enrichment.ratio.se*1.96),width=0.01)



############### release summary
enrichment.conc<-aggregate(conservative.NP[,2]~NP,data=release,mean)
enrichment.conc.se<-aggregate(conservative.NP[,2]~NP,data=release,se)

uptake.ratio<-aggregate(uptake.NP[,2]~NP,data=release,mean)
uptake.ratio.se<-aggregate(uptake.NP[,2]~NP,data=release,se)



release.summary<-data.frame(NP=enrichment.conc[,1],mean.enrichment.ratio=enrichment.conc[,2],enrichment.ratio.se=enrichment.conc.se[,2],uptake.ratio.mean=uptake.ratio[,2],uptake.ratio.se=uptake.ratio.se[,2])

release$release.NP<-release$uptake.NP[,2]
release$conservative.NP.plot<-release$conservative.NP[,2]

log.normal.plot.release<-ggplot(release,aes(x=conservative.NP.plot,y=release.NP,color=as.factor(NP),shape=as.factor(NP)))+geom_abline(intercept=0.45,slope=0.823,size=1,color="black")+ theme_classic()+ geom_abline(intercept=0,slope=1,size=1,linetype="dashed")+theme(text =element_text(size=20))+ scale_x_log10()+scale_y_log10()+scale_color_manual(values=cbbPalette)+xlab("Added + Background N:P ratio")+ylab("Net release N:P ratio")+scale_shape_manual(values=c(0, 1, 2,18,19))+theme(legend.position = "none")+xlim(1,150)+ylim(1,150)+geom_point(size=5)#+geom_errorbar(aes(ymin=uptake.ratio.mean-uptake.ratio.se*1.96,ymax=uptake.ratio.mean+uptake.ratio.se*1.96),width=0.01)+geom_errorbarh(aes(xmin=mean.enrichment.ratio-enrichment.ratio.se*1.96,xmax=mean.enrichment.ratio+enrichment.ratio.se*1.96),width=0.01)

log.normal.plot.release


### Simple plot
N.averages<-aggregate(n.uptake.uM.m2.sec~date+as.factor(NP),compiled_data,median)
P.averages<-aggregate(p.uptake.uM.m2.sec~date+as.factor(NP),compiled_data,median)

averages<-merge(N.averages,P.averages,by=c("date","as.factor(NP)"))

averages$NP<-averages$`as.factor(NP)`

simple.plot<-ggplot(averages,aes(y=n.uptake.uM.m2.sec,x=p.uptake.uM.m2.sec,color=NP,shape=NP))+geom_point(size=2)+geom_smooth(method="lm",se=FALSE)+theme_classic()+xlab(expression(italic("NNE")~"SRP ("*mu*mol~m^-2~s^-1*")"))+ylab(expression(italic("NNE")~"DIN ("*mu*mol~m^-2~s^-1*")"))+scale_color_manual(values=cbbPalette,name="")+theme(text =element_text(size=20))+theme(legend.position = "none")+geom_abline(intercept = 0,slope=16,linetype="dashed")+ scale_shape_manual(values=c(0, 1, 2,18,19))
simple.plot

tiff(filename="new.stoich.plot.feb_2022.tiff",units="in",res=600,width=8,height=16,compression="lzw")
plot_grid(simple.plot,log.normal.plot,log.normal.plot.release,ncol=1,labels="AUTO",label_x=0.85,label_y=0.98,label_size=20)
dev.off()

simple.model<-lm(n.uptake.uM.m2.sec~p.uptake.uM.m2.sec*NP,data=averages)
anova(simple.model)
summary(simple.model)


tiff(filename="new.stoich.plot.jan_2022.tiff",units="in",res=600,width=6,height=10,compression="lzw")
plot_grid(simple.plot,log.normal.plot,ncol=1,labels="AUTO",label_x=0.9,label_y=0.25)
dev.off()




#what are the slopes of each line  ## These presented in table A3
summary(lm(n.uptake.uM.m2.sec~p.uptake.uM.m2.sec,data=averages[averages$NP==2,]))
summary(lm(n.uptake.uM.m2.sec~p.uptake.uM.m2.sec,data=averages[averages$NP==8,]))
summary(lm(n.uptake.uM.m2.sec~p.uptake.uM.m2.sec,data=averages[averages$NP==16,]))
summary(lm(n.uptake.uM.m2.sec~p.uptake.uM.m2.sec,data=averages[averages$NP==32,]))
summary(lm(n.uptake.uM.m2.sec~p.uptake.uM.m2.sec,data=averages[averages$NP==128,]))


## what is the log-log slope of inputs vs. uptake (calculating H-1)
log.linear<-lm(log(uptake.NP[,2])~log(conservative.NP[,2]),data=stoich2)
anova(log.linear)
summary(log.linear)

## is the log-log slope different than 1?
log.linear<-lm(log(uptake.NP[,2])~log(conservative.NP[,2]),offset= 1.00*log(conservative.NP[,2]),data=stoich2)
anova(log.linear)
summary(log.linear)

####### What about release
log.release<-lm(log(uptake.NP[,2])~log(conservative.NP[,2]),data=release)
summary(log.release)
anova(log.release)

log.release.2<-lm(log(uptake.NP[,2])~log(conservative.NP[,2]),offset= 1.00*log(conservative.NP[,2]),data=release)
summary(log.release.2)
anova(log.release.2)

```

writing file for further analysis
```{r}

write.csv(compiled_data,file="Final_uptake_estiamtes.csv")

write.csv(P.quants,file="P_uptake_status.csv")
write.csv(N.quants,"N.uptake.status.csv")
```

# Do the residiuals in figure 5a have a realtionship to time

```{r}
averages$residuals<-residuals(simple.model)

residuals.time.model<-lm(residuals~date*as.numeric(as.character(NP)),averages)
anova(residuals.time.model)

summary(residuals.time.model)

residuals.time<-ggplot(averages,aes(y=residuals,x=date,color=NP,shape=NP))+geom_point(size=2)+geom_smooth(method="lm",se=FALSE)+theme_classic()+xlab(expression("Date"))+ylab(expression("Stoichiometry residuals"))+scale_color_manual(values=cbbPalette,name="")+theme(text =element_text(size=20))+theme(legend.position = "none")+scale_shape_manual(values=c(0, 1, 2,18,19))
residuals.time

## going to print this later

```

